---
title: "Hatch Rate Analysis 2016"
output:
  pdf_document: default
  html_document: default
date: "June 19, 2017"
---
Setup
```{r}
library(knitr)
Sys.setenv(RSTUDIO_PDFLATEX = "/Library/TeX/texbin/pdfLaTex")
```

Function that allows line wrapping of long code
```{r}
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```

Set up Environment:
```{r}
library('reshape2')
library('scales')
library('stringr')
library('ggplot2')
setwd('/Users/Tessa/Dropbox/squid/squid_R_work/squid_hatching/hatching_rmd')
source('helpful_summary_functions.R') #load helper functions
hr <-read.csv('hatch_rate_transposed_updated.csv', header=T)
hr$Date <- as.Date(hr$Date, "%m/%d/%y") # format Date properly
```

##Cumulative hatching over time:

Calulate cumulative hatching over time, per capsule ('2' = by column):
```{r}
cumSumHR <- apply(hr[,-which(names(hr) == "Date")],2,cumsum) #don't sum date column
```

Join to original dataframe so date column can correspond
```{r}
#add 'cumul' to column names to diff from orig count info
colnames(cumSumHR) <- paste(colnames(cumSumHR), 'cumul', sep = "_") 
#join DF's
fullHR_DF <- cbind(hr,cumSumHR)
#could add dates here if we want
#cumSumHR_wDates <- fullHR_DF[, grepl("cumul", colnames(fullHR_DF))]
#cumSumHR_wDates$Date <- fullHR_DF$Date
```

Melt to long-form; add classifying variables
```{r}
meltedFullDF <- melt(fullHR_DF, id='Date')
meltedFullDF$Tank <- ""
meltedFullDF[grepl("1", meltedFullDF$variable),]$Tank <- 'Tank 1'
meltedFullDF[grepl("2", meltedFullDF$variable),]$Tank <- 'Tank 2'
meltedFullDF[grepl("3", meltedFullDF$variable),]$Tank <- 'Tank 3'
meltedFullDF[grepl("4", meltedFullDF$variable),]$Tank <- 'Tank 4'
meltedFullDF$Condition <- ""
meltedFullDF[grepl("1|2", meltedFullDF$variable),]$Condition <- 'pH 7.57, 97 uM/kg O2'
meltedFullDF[grepl("3|4", meltedFullDF$variable),]$Condition <- 'pH 7.89, 227 uM/kg O2'
```

Split cumulative counts; original counts --> two DF's
```{r}
melted_CumOnly <- meltedFullDF[grepl("cumul", meltedFullDF$variable),]
melted_countsOnly <- meltedFullDF[!grepl("cumul", meltedFullDF$variable),]
```

Summarize cumulative count data; get Date in appropriate form for graphing
```{r}
summCond_Cumul <- summarySE(melted_CumOnly, measurevar="value", groupvars=c("Condition","Date"))
summCond_Cumul$Date <- as.Date(summCond_Cumul$Date)
```

Read in unhatched count data; test for diffs btwn treatments
```{r}
unhatched_embs <- read.csv('unhatched_embryos.csv', header=T)

#unhatchMelt <- melt(unhatched_embs, id="Date")
#unhatchMelt$Tank <- ""
#unhatchMelt[grepl("1", unhatchMelt$variable),]$Tank <- 'Tank 1'
#unhatchMelt[grepl("2", unhatchMelt$variable),]$Tank <- 'Tank 2'
#unhatchMelt[grepl("3", unhatchMelt$variable),]$Tank <- 'Tank 3'
#unhatchMelt[grepl("4", unhatchMelt$variable),]$Tank <- 'Tank 4'
#unhatchMelt$Condition <- ""
#unhatchMelt[grepl("1|2", unhatchMelt$variable),]$Condition <- 'pH 7.57, 97 uM/kg O2'
#unhatchMelt[grepl("3|4", unhatchMelt$variable),]$Condition <- 'pH 7.89, 227 uM/kg O2'

#unhatchSum_cond <- summarySE(unhatchMelt, measurevar="value", groupvars=c("Condition"))
#unhatchMelt$Condition <- as.factor(unhatchMelt$Condition)

#Get total embs per capsule
totalHatched = colSums(hr[,-1])
totalUnhatched =unhatched_embs[,-1]
embs_per_cap = totalHatched + totalUnhatched
```

Is %hatch diff btwn treatments?
```{r}
percentHatch = (totalUnhatched/embs_per_cap)*100
meltedPhatch = melt(percentHatch)
meltedPhatch$Tank <- ""
meltedPhatch[grepl("1", meltedPhatch$variable),]$Tank <- 'Tank 1'
meltedPhatch[grepl("2", meltedPhatch$variable),]$Tank <- 'Tank 2'
meltedPhatch[grepl("3", meltedPhatch$variable),]$Tank <- 'Tank 3'
meltedPhatch[grepl("4", meltedPhatch$variable),]$Tank <- 'Tank 4'
meltedPhatch$Condition <- ""
meltedPhatch[grepl("1|2", meltedPhatch$variable),]$Condition <- 'pH 7.57, 97 uM/kg O2'
meltedPhatch[grepl("3|4", meltedPhatch$variable),]$Condition <- 'pH 7.89, 227 uM/kg O2'
meltedPhatch$percentHatched <- 100 - meltedPhatch$value
pHatched_treat =meltedPhatch[grepl("1|2", meltedPhatch$variable),]$percentHatched
pHatched_ctrl =meltedPhatch[grepl("3|4", meltedPhatch$variable),]$percentHatched
t.test(pHatched_ctrl, pHatched_treat, alternative="greater")
```

##Plot percent hatching over time

first, get total embs per cap into the cumulative dataframe
```{r}
#cap name --> column
melted_CumOnly$Capsule <- str_match(melted_CumOnly$variable,'(X\\d\\S)_cumul')[,2] 
#get # per capsule into right form
embs_per_capMelt <- melt(embs_per_cap)
names(embs_per_capMelt) <- c('Capsule', 'embTotal')
#merge the two dataframes by intersecting the capsule name columns
melted_CumOnly_wTotals = merge(melted_CumOnly, embs_per_capMelt, by = intersect("Capsule", "Capsule"))
melted_CumOnly_wTotals$percentTotalHatched = melted_CumOnly_wTotals$value/melted_CumOnly_wTotals$embTotal
```

Get summary of % Hatching
```{r}
summCond_PercentCumul <- summarySE(melted_CumOnly_wTotals,    measurevar="percentTotalHatched", groupvars=c("Condition","Date"))
summCond_PercentCumul$Date <- as.Date(summCond_PercentCumul$Date)
```


##Include incubation time; x-axis --> days, instead of 'Date'
```{r,tidy=TRUE}
#add in inc time:
inc_dates <- seq.Date(as.Date('2013/06/02'), as.Date('2013/07/02'), by='days')
zero_hatching <- rep(0,length(inc_dates))
zero_hatch_df <- as.data.frame(matrix(zero_hatching,nrow = length(zero_hatching),ncol = 16))
inc_only <-  cbind(inc_dates, zero_hatch_df)
colnames(inc_only) <- colnames(hr)
hr_with_incubation <- merge(hr,inc_only, on='Date', all=T)
hr_with_incubation$days = hr_with_incubation$Date - as.Date('2013/06/02')

cumSumHR_w_inc <- apply(hr_with_incubation[,-which(names(hr_with_incubation) == c("Date", "days"))],2,cumsum)
colnames(cumSumHR_w_inc) <- paste(colnames(cumSumHR_w_inc), 'cumul', sep = "_")

fullHR_DF_w_inc <- cbind(hr_with_incubation,cumSumHR_w_inc)
meltedFullDF_w_inc <- melt(fullHR_DF_w_inc, id=c('Date', 'days'))
meltedFullDF_w_inc$Tank <- ""
meltedFullDF_w_inc[grepl("1", meltedFullDF_w_inc$variable),]$Tank <- 'Tank 1'
meltedFullDF_w_inc[grepl("2", meltedFullDF_w_inc$variable),]$Tank <- 'Tank 2'
meltedFullDF_w_inc[grepl("3", meltedFullDF_w_inc$variable),]$Tank <- 'Tank 3'
meltedFullDF_w_inc[grepl("4", meltedFullDF_w_inc$variable),]$Tank <- 'Tank 4'
meltedFullDF_w_inc$Condition <- ""
meltedFullDF_w_inc[grepl("1|2", meltedFullDF_w_inc$variable),]$Condition <- 'pH 7.57, 97 uM/kg O2'
meltedFullDF_w_inc[grepl("3|4", meltedFullDF_w_inc$variable),]$Condition <- 'pH 7.89, 227 uM/kg O2'

melted_CumOnly_w_inc <- meltedFullDF_w_inc[grepl("cumul", meltedFullDF_w_inc$variable),]
melted_countsOnly_w_inc <- meltedFullDF_w_inc[!grepl("cumul", meltedFullDF_w_inc$variable),]

melted_CumOnly_w_inc$Capsule <- str_match(melted_CumOnly_w_inc$variable,'(X\\d\\S)_cumul')[,2] 
melted_CumOnly_w_incT = merge(melted_CumOnly_w_inc, embs_per_capMelt, by = intersect("Capsule", "Capsule"))
melted_CumOnly_w_incT$percentTotalHatched = melted_CumOnly_w_incT$value/melted_CumOnly_w_incT$embTotal


summCond_PercentCumul_w_incT <- summarySE(melted_CumOnly_w_incT, measurevar="percentTotalHatched", groupvars=c("Condition","days"))


#pdf(file="percent_hatch_by_condition_w100line_w_incTime_brcolor.pdf",width=7,height=4.5)
#pdf(file="percent_hatch_by_condition_w90phatchlines_wincTime_brcolor.pdf",width=7,height=4.5)
#pdf(file="percent_hatch_by_condition_w90phatchlines_wincTime_brcolor.pdf",width=7,height=4.5)
#pdf(file="percent_hatch_by_condition_w90phatchlines_wincTime_color_notreatbox_realtime.pdf",width=7,height=4.5)
ggplot(summCond_PercentCumul_w_incT, aes(x=days, y=percentTotalHatched, shape=Condition, color=Condition)) + geom_errorbar(aes(ymin=percentTotalHatched-se, ymax=percentTotalHatched+se), color='darkgrey', width=.05) + geom_line() + geom_point(size = 3)  + scale_color_manual(values=c("#d7191c", "black"))+ xlab('Days since egg deposition') + ylab("Percent Hatched") + ggtitle(expression(paste("Cumulative Percent Hatching of ",italic("Doryteuthis opalescens"), " Embryos")))  + theme_bw() + theme(plot.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank() )+ theme(panel.border= element_blank()) + theme(legend.key = element_blank(),legend.justification=c(.05,0.85), legend.position=c(.15,0.65)) + geom_rect(xmin=0, xmax=2, ymin=-Inf, ymax=+Inf, fill='gray75', alpha=0.01,inherit.aes = FALSE) + geom_vline(xintercept=39, linetype='dotted') + geom_vline(xintercept=43.5, linetype='dotted', color="#e41a1c") + theme(axis.line.x = element_line(color="black", size = 1), axis.line.y = element_line(color="black", size = 1))
#dev.off()



# for 90% of each total Hatched () 90% of 99.8% and 90% of 94.5%): add horizontal line at 90 or 100 + geom_hline(yintercept=.9, linetype='dotted')
# doesn't look great. instead, let's try adding vertical lines at incubation time = 90% hatch (39d, 43.5d)
# + geom_vline(xintercept=39, linetype='dotted') + geom_vline(xintercept=43.5, linetype='dotted')

#try adding shaded rect for the time we didn't have embryos:
#+ geom_rect(xmin=0, xmax=3, ymin=-Inf, ymax=+Inf), fill='grey', alpha=0.2)

melted_CumOnly_w_incT[grepl("1|2", melted_CumOnly_w_incT$variable),]$Condition <- 'pHOx'
melted_CumOnly_w_incT[grepl("3|4", melted_CumOnly_w_incT$variable),]$Condition <- 'control'

summTank_PercentCumul_w_incT <- summarySE(melted_CumOnly_w_incT, measurevar="percentTotalHatched", groupvars=c("Tank","days"))
```

Calculate # for 90% hatching (see Kaplan et al)
```{r}
hatch90percent = data.frame((.9*totalHatched))
colnames(hatch90percent) <- c('hatch90percent')
hatch90percent$Capsule <- rownames(hatch90percent)
```

Is incubation time diff btwn treatments?
```{r}

melted_CumOnly_w90p = merge(melted_CumOnly,hatch90percent, by = intersect("Capsule", "Capsule"))
over90phatched <- melted_CumOnly_w90p[melted_CumOnly_w90p$value >= melted_CumOnly_w90p$hatch90percent, ]

incubation90<- ddply(over90phatched, .(Capsule), summarize, end_incubation=min(Date))
incubation90$inc_start <- as.Date('2013-06-01')
incubation90$days <- incubation90$end_incubation - incubation90$inc_start
#add metadata
incubation90$Tank <- ""
incubation90[grepl("1", incubation90$Capsule),]$Tank <- 'Tank 1'
incubation90[grepl("2", incubation90$Capsule),]$Tank <- 'Tank 2'
incubation90[grepl("3", incubation90$Capsule),]$Tank <- 'Tank 3'
incubation90[grepl("4", incubation90$Capsule),]$Tank <- 'Tank 4'
incubation90$Condition <- ""
incubation90[grepl("1|2", incubation90$Capsule),]$Condition <- 'pH 7.57, 97 uM/kg O2'
incubation90[grepl("3|4", incubation90$Capsule),]$Condition <- 'pH 7.89, 227 uM/kg O2'

#t-test
inc90_treat =incubation90[grepl("1|2", incubation90$Capsule),]$days
inc90_ctrl =incubation90[grepl("3|4", incubation90$Capsule),]$days
t.test(inc90_ctrl, inc90_treat, alternative="less")
```

SE of mean of incubation90 values
```{r}
se <- function(x) sqrt(var(x)/length(x))
se_inc90_ctrl <- se(inc90_ctrl)
se_inc90_treat <- se(inc90_treat)
mean_inc90_ctrl <- mean(inc90_ctrl)
mean_inc90_treat <- mean(inc90_treat)
```

##Duration of hatching (reach 90% hatching as incubation time)

```{r}
incubation90$hatch_start <- as.Date('2013-07-03')
incubation90[grepl("1|2", incubation90$Capsule),]$hatch_start <- as.Date('2013-07-06')
incubation90$hatching_duration <- incubation90$end_incubation - incubation90$hatch_start

hdur90_treat =incubation90[grepl("1|2", incubation90$Capsule),]$hatching_duration
hdur90_ctrl =incubation90[grepl("3|4", incubation90$Capsule),]$hatching_duration
t.test(hdur90_ctrl, hdur90_treat, alternative="less")
```


#  Chi-Sq to compare two distributions:
http://thebiobucket.blogspot.com/2011/08/comparing-two-distributions.html#more

As in Kaplan et al, test if the distributions of hatching over time is the same for all treatments.
Use COUNT data:
```{r}
hr_for_ch <- hr
hr_for_ch$lowpHOx <- rowSums(hr_for_ch[, grepl('1|2',colnames(hr_for_ch))])
hr_for_ch$highpHOx <- rowSums(hr_for_ch[, grepl('3|4',colnames(hr_for_ch))])

chtable <- hr_for_ch[, grepl('pHOx', colnames(hr_for_ch))]
#chisq.test(chtable)

chisq.test(t(chtable)) # treatments need to be rows
```

Check temperature ~ percent hatching:
```{r}
pHatched_13.5 =meltedPhatch[grepl("1|3", meltedPhatch$variable),]$percentHatched
pHatched_12 =meltedPhatch[grepl("2|4", meltedPhatch$variable),]$percentHatched
t.test(pHatched_12, pHatched_13.5, alternative = 'less')
```
```


